Measurement of total hadronic differential cross sections in the LArIAT experiment

### 2018-05-21 Chapter-4/4-Methodology.tex
s

As we will show in section \ref{sec:Correction}, the background subtraction for the interacting and incident histograms can be translated into a corresponding relative pion content $C^{\pi MC}_{Interacting} (E_{i})$ and $C^{\pi MC}_{Incident} (E_{i})$ and the cross section re-written as follows

>>>>

As we will show in section \ref{ch:PionXSBkgSub}, the background subtraction for the interacting and incident histograms can be translated into a corresponding relative pion content $C^{\pi MC}_{Interacting} (E_{i})$ and $C^{\pi MC}_{Incident} (E_{i})$ and the cross section re-written as follows

### 2018-05-21 Chapter-5/5-Preparatory.tex
s

\begin{figure}[p]
\centering
\includegraphics[width=15cm]{Chapter-7/Images/CDAsMomentumFunct.png}
\caption{True momentum distribution at wire chamber 4 for every simulated pion arriving in the TPC (pink), ending its life in capture (green) or in decay (blue) in the TPC, linear vertical axis on the left, logarithmic on the right. }
\label{fig:CaptureMom}
\end{figure}

\begin{figure}[p]
\centering
\begin{minipage}[t]{0.45\textwidth}
\centering
\includegraphics[width=7.5cm]{Chapter-7/Images/CDThreshold.png}
\caption{Survival ratio as a function of selection threshold on true momentum at wire chamber four for for every simulated pion arriving in the TPC (pink), capture (green) or in decay (blue).   }
\label{fig:survRatio}
\end{minipage}\hfill
\begin{minipage}[t]{0.45\textwidth}
\centering
\includegraphics[width=7.5cm]{Chapter-7/Images/CDRatio.png}
\caption{Ratio between the capture (green) and decay (blue) events over the total number of events as a as a function of the true momentum at wire chamber four.}
\label{fig:evtRatio}
\end{minipage}
\end{figure}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Beamline Background}\label{ch:PionXSBkgSub2}
We define beamline background every TPC track matched to the WC track which is not a primary pion. Potentially, there are 4 different types of beamline background:
\begin{itemize}
\item[]1) electrons,
\item[]2) muons,
\item[]3) secondaries from pion events,
\item[]4) matched pile up events.
\end{itemize}

The first step is to estimate what percentage of events used in the cross section calculation is not a primary pion.  We start by noting that the last type of background, the ``matched pile up" events, is a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} negligibly small. %\textcolor{red}{SHOW VTX distribution in WC2TPC match}

As shown in Section \ref{ch:beamlineComposition}, we use G4Beamline to estimate the beam composition at WC4, obtaining the composition shown in Table \ref{tab:beamline}.
The next step to estimate the beamline background in the cross section measurement is propagating pions, muons and electrons to the TPC and evaluate their contribution to the cross section. To do so, we simulate the same number of electrons, muons and pions with the DDMC and we apply the same selection filters on the three samples. The number of events per particle species surviving this selection is shown on table \ref{tab:MCafterCutContaminants}.

Pions can travel the length of the LArIAT beamline and interact hadronically in the steel or in the non-instrumented argon upstream to the TPC front face. Or, they could decay in flight between WC4 and the TPC. One of the interaction products can leak into the TPC and be matched with the WC track, contributing to the pool of events used for the cross section calculation. We call this type of particles ``secondaries" from pion events, with a terminology inspired by Geant4. 
We estimate the number of secondaries using the DDMC pion sample.  The percentage of secondaries is given by the number of matched WC2TPC tracks whose corresponding particle is not flagged as primary by Geant4.  The secondary to pion ratio is 4.9\% in the 60A sample and 4.3\% in the 100A sample.

@@

In order to reproduce the closest make up of the beam to data, we weight each event of a given particle species according to the estimated beam composition. In case of 60A runs, for example, the weights are 0.688 for pions,  0.046 for muons  and 0.266 for electrons.
We produce accordingly the interacting and incident histograms for the events surviving the selection, staggering the contributions for each particle species, as shown in Figure  \ref{fig:stag60A}. From those histograms, we are able to evaluate the relative contribution of  pions and  background to each bin of the interacting and incident histograms separately and obtain the respective corrections for data. We take here the interacting histogram as example, noting that the derivation of the correction for the incident histogram is identical. The number of entries in each bin of the interacting plot (Figure \ref{fig:stag60A} left) is  $N^{\text{TOT}}_{\text{Interacting}} (E_{i})$, equal to the sum of the pions and background in that bin, namely

\begin{equation}
N^{\text{TOT}}_{\text{Interacting}} (E_{i}) =  N^\pi_{Interacting} (E_{i}) + \underbrace{ N^\mu_{Interacting} (E_{i}) + N^e_{Interacting} (E_{i}) + N^{Secondary}_{Interacting} (E_{i}) }_{B_{Interacting} (E_i)}.
\end{equation}
Thus, the relative contribution of pions to each bin in MC can be calculated as follows
\begin{equation}
C^{\pi MC}_{Interacting} (E_{i}) =  \frac{N^{\pi MC}_{Interacting}}{ N^{TOT MC}_{Interacting} (E_{i}) } =    \frac{N^{TOT MC}_{Interacting} (E_{i}) - B^ {MC}_{Interacting} (E_i)}{ N^{TOT MC}_{Interacting} (E_{i})}.
\end{equation}

In order to evaluate the pion content of each bin in data, we scale the measured bin content by the corresponding  pion contribution found in MC, as follows
\begin{equation}
N^{\pi Reco Data}_{Interacting} = N^{TOT Data}_{Interacting} (E_{i}) - B^{Data}_{Interacting} (E_i)  =  C^{\pi MC}_{Interacting} (E_{i}) N^{TOT Data}_{Interacting} (E_{i}).
\end{equation}

>>>>

\begin{figure}[]
\centering
\includegraphics[width=15cm]{Chapter-7/Images/CDAsMomentumFunct.png}
\caption{True momentum distribution at wire chamber 4 for every simulated pion arriving in the TPC (pink), ending its life in capture (green) or in decay (blue) in the TPC, linear vertical axis on the left, logarithmic on the right. }
\label{fig:CaptureMom}
\end{figure}

\begin{figure}[]
\centering
\begin{minipage}[t]{0.45\textwidth}
\centering
\includegraphics[width=7.5cm]{Chapter-7/Images/CDThreshold.png}
\caption{Survival ratio as a function of selection threshold on true momentum at wire chamber four for for every simulated pion arriving in the TPC (pink), capture (green) or in decay (blue).   }
\label{fig:survRatio}
\end{minipage}\hfill
\begin{minipage}[t]{0.45\textwidth}
\centering
\includegraphics[width=7.5cm]{Chapter-7/Images/CDRatio.png}
\caption{Ratio between the capture (green) and decay (blue) events over the total number of events as a as a function of the true momentum at wire chamber four.}
\label{fig:evtRatio}
\end{minipage}
\end{figure}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Contributions from the Beamline Background}\label{ch:PionXSBkgSub2}
We define beamline background every TPC track matched to the WC track which is not a primary pion. Potentially, there are 4 different types of beamline background:
\begin{itemize}
\item[]1) electrons,
\item[]2) muons,
\item[]3) secondaries from pion events,
\item[]4) matched pile up events.
\end{itemize}

The first step to quantify the effect of the beamline background on the pion cross section is to estimate what percentage of events used in the cross section calculation is not a primary pion.  We start by noting that the last type of background, the ``matched pile up" events, is a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} negligibly small. %\textcolor{red}{SHOW VTX distribution in WC2TPC match}
As shown in Section \ref{ch:beamlineComposition}, we use G4Beamline to estimate the percentage of pions, muons and electrons at WC4, obtaining the composition shown in Table \ref{tab:beamline}. The next step is to simulate those pions, muons and electrons from WC4 to the TPC with the DDMC and evaluate their contribution to the cross section. To do so, we start by simulating the same number of electrons, muons and pions with the DDMC and we apply the same selection filters on the three samples. The number of events per particle species surviving this selection is shown on table \ref{tab:MCafterCutContaminants}. In order to reproduce the closest make up of the beam to data, we weight each event of a given particle species according to the estimated beam composition. In case of 60A runs, for example, the weights are 0.688 for pions,  0.046 for muons  and 0.266 for electrons.

It should be noted that pions may  interact hadronically in the steel or in the non-instrumented argon upstream to the TPC front face while travelling the length of between WC4 and the TPC. Or, they could decay in flight between WC4 and the TPC. One of the interaction products can leak into the TPC and be matched with the WC track, contributing to the pool of events used for the cross section calculation. We call this type of particles ``secondaries" from pion events, with a terminology inspired by Geant4.  We estimate the number of secondaries using the DDMC pion sample.  The percentage of secondaries is given by the number of matched WC2TPC tracks whose corresponding particle is not flagged as primary by Geant4.  The secondary to pion ratio is 4.9\% in the 60A sample and 4.3\% in the 100A sample.

@@

We evaluate the beamline background contribution to the cross section by producing the interacting and incident histograms for the events surviving the selection, staggering the contributions for each particle species, as shown in Figure  \ref{fig:stag60A}. From those histograms, we are able to evaluate the contribution of  pions and  beamline backgrounds to each bin of the interacting and incident histograms separately and obtain the relative pion content. The relative pion content in each bin for the interacting and incident histograms represents the correction applied to data. We take here the interacting histogram as example, noting that the derivation of the correction for the incident histogram is identical. The number of entries in each bin of the interacting plot (Figure \ref{fig:stag60A} left) is  $N^{\text{TOT}}_{\text{Int}} (E_{i})$, equal to the sum of the pions and beamline backgrounds in that bin, namely

\begin{equation}
N^{\text{TOT}}_{\text{Int}} (E_{i}) =  N^\pi_{\text{Int}} (E_{i}) + \underbrace{ N^\mu_{\text{Int}} (E_{i}) + N^e_{\text{Int}} (E_{i}) + N^{Secondary}_{\text{Int}} (E_{i}) }_{B_{\text{Int}} (E_i)}.
\end{equation}
Thus, the relative pion content to each bin in MC can be calculated as follows
\begin{equation}
C^{\pi MC}_{\text{Int}} (E_{i}) =  \frac{N^{\pi MC}_{\text{Int}}}{ N^{TOT MC}_{\text{Int}} (E_{i}) } =    \frac{N^{TOT MC}_{\text{Int}} (E_{i}) - B^ {MC}_{\text{Int}} (E_i)}{ N^{TOT MC}_{\text{Int}} (E_{i})}.
\end{equation}

In order to evaluate the pion content of each bin in data, we scale the measured bin by the corresponding relative pion content found in MC, as follows
\begin{equation}
N^{\pi Reco Data}_{\text{Int}} = N^{TOT Data}_{\text{Int}} (E_{i}) - B^{Data}_{\text{Int}} (E_i)  =  C^{\pi MC}_{\text{Int}} (E_{i}) N^{TOT Data}_{\text{Int}} (E_{i}).
\end{equation}

The pion content is evaluated separately in the interacting and incident histograms. Their ratio determines a correction to the measured raw cross section. 
For example, the measured raw  cross section of a sample with enhanced muons content will tend to be lower than the raw cross section of a muon free sample. This is because most of the muons will cross the TPC without stopping, thus contributing almost exclusively to the incident histogram, forcing the pion content to be lower in the incident histogram than in the interacting; thus, the correction will tend to enhance the cross section.

### 2018-05-21 Chapter-6/6-PionCrossSection.tex
s

Section \label{ch:BKGsubXS} describes the procedure employed to obtain  
$C^{\pi MC}_{\text{Int}} (E_{i})$  and  $C^{\pi MC}_{\text{Inc}} (E_{i})$ and the propagation to the cross section measurement of the relative uncertainties.

Section \label{ch:ch:EFFXS} describes the procedure employed to obtain  
$\epsilon^{\text{Int}}(E_i)$  and $\epsilon^{\text{Inc}}(E_i)$ and the propagation to the cross section measurement of the relative uncertainties.

\subsection{Background subtraction}\label{ch:BKGsubXS}
%Even if pions are by far the biggest component of the beam in negative polarity runs, the LArIAT beam is not a pure pion beam. While useful to discriminate  pions/muons/electrons from kaons, and protons, the beamline detectors are not sensitive enough to  discriminate among the lighter particles in the beam: electrons, muons and pions fall under the same mass hypothesis. Thus, we need to assess the background from beamline particles other than pions in the event selections used for the pion cross section analysis and correct for its effects.

%\subsubsection{Background from secondaries at TPC Front Face}
%Pions can travel the length of the LArIAT beamline and interact hadronically in the steel or in the non-instrumented argon upstream to the TPC front face. Or, they could decay in flight between WC4 and the TPC. One of the interaction products can leak into the TPC and be matched with the WC track, contributing to the pool of events used for the cross section calculation. We call this type of particles ``secondaries" from pion events, with a terminology inspired by Geant4. 
We estimate the number of secondaries using the DDMC pion sample.  The percentage of secondaries is given by the number of matched WC2TPC tracks whose corresponding particle is not flagged as primary by Geant4.  The secondary to pion ratio is 4.9\% in the 60A sample and $Y$\% in the 100A sample.

@@

>>>>

Section \ref{ch:BKGsubXSPuppa} describes the evaluation of pion content in the interacting and incident histograms, ($C^{\pi MC}_{\text{Int}} (E_{i})$  and  $C^{\pi MC}_{\text{Inc}} (E_{i})$) and the propagation to the cross section measurement of the relative systematic uncertainties.

Section \ref{ch:EFFXS} describes the procedure employed to obtain  the efficiency corrections $\epsilon^{\text{Int}}(E_i)$  and $\epsilon^{\text{Inc}}(E_i)$ and the propagation to the cross section measurement of the relative uncertainties.

\subsection{Background subtraction}\label{ch:BKGsubXSPuppa}
We use the procedure described in \ref{ch:PionXSBkgSub2} to evaluate the pion content in the interacting $C^{\pi MC}_{\text{Int}} (E_{i})$  and incident $C^{\pi MC}_{\text{Inc}} (E_{i})$) histograms.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{comment}
\subsection{Background subtraction}\label{ch:BKGsubXS}
%Even if pions are by far the biggest component of the beam in negative polarity runs, the LArIAT beam is not a pure pion beam. While useful to discriminate  pions/muons/electrons from kaons, and protons, the beamline detectors are not sensitive enough to  discriminate among the lighter particles in the beam: electrons, muons and pions fall under the same mass hypothesis. Thus, we need to assess the background from beamline particles other than pions in the event selections used for the pion cross section analysis and correct for its effects.

%\subsubsection{Background from secondaries at TPC Front Face}
%Pions can travel the length of the LArIAT beamline and interact hadronically in the steel or in the non-instrumented argon upstream to the TPC front face. Or, they could decay in flight between WC4 and the TPC. One of the interaction products can leak into the TPC and be matched with the WC track, contributing to the pool of events used for the cross section calculation. We call this type of particles ``secondaries" from pion events, with a terminology inspired by Geant4. 
%We estimate the number of secondaries using the DDMC pion sample.  The percentage of secondaries is given by the number of matched WC2TPC tracks whose corresponding particle is not flagged as primary by Geant4.  The secondary to pion ratio is 4.9\% in the 60A sample and $Y$\% in the 100A sample.

\end{comment}
